% Database Structure - Tables and constraints
\section{Database Structure}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\subsection{User Login Table (\texttt{user\_login})}

\begin{tabularx}{\textwidth}{@{}L{3.5cm} L{1.5cm} L{4cm} X@{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Constraints} & \textbf{Notes} \\
\midrule
\texttt{user\_id} & INTEGER & PRIMARY KEY AUTOINCREMENT & Unique user identifier \\
\texttt{username} & TEXT & UNIQUE, NOT NULL & Indexed for lookup \\
\texttt{email} & TEXT & UNIQUE, NOT NULL & Stored normalized (lowercase) \\
\texttt{password} & TEXT & NOT NULL & Argon2 hash in PHC format \\
\texttt{email\_verified} & INTEGER & NOT NULL, DEFAULT 0 & Boolean: 0=false, 1=true \\
\texttt{email\_verified\_at} & INTEGER & NULLABLE & Unix timestamp when verified \\
% Login_4 scope - password reset field
% \texttt{password\_reset} & INTEGER & NOT NULL, DEFAULT 0 & Boolean: password reset in progress \\
\bottomrule
\end{tabularx}

\vspace{0.5em}
Constraints and notes:
\begin{itemize}[leftmargin=*,nosep]
  \item Unique constraints on \texttt{username} and \texttt{email} prevent duplicates.
  \item \texttt{email\_verified} must be true before user can log in.
\end{itemize}

\subsection{User Data Table (\texttt{user\_data})}

\begin{tabularx}{\textwidth}{@{}L{3.5cm} L{1.5cm} L{4cm} X@{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Constraints} & \textbf{Notes} \\
\midrule
\texttt{user\_id} & INTEGER & PRIMARY KEY, FOREIGN KEY $\rightarrow$ \texttt{user\_login.user\_id} & CASCADE on delete \\
\texttt{counter} & INTEGER & NOT NULL, DEFAULT 0 & Application-specific data \\
\bottomrule
\end{tabularx}

\vspace{0.5em}
Notes:
\begin{itemize}[leftmargin=*,nosep]
  \item One-to-one relationship with \texttt{user\_login}.
  \item Created automatically during registration.
  \item Cascading delete ensures cleanup when user is removed.
\end{itemize}

\subsection{Email Verification Tokens Table (\texttt{email\_verification\_tokens})}

\begin{tabularx}{\textwidth}{@{}L{3.5cm} L{1.5cm} L{4cm} X@{}}
\toprule
\textbf{Column} & \textbf{Type} & \textbf{Constraints} & \textbf{Notes} \\
\midrule
\texttt{user\_id} & INTEGER & PRIMARY KEY, FOREIGN KEY $\rightarrow$ \texttt{user\_login.user\_id} & CASCADE on delete \\
\texttt{token\_hash} & TEXT & NOT NULL & SHA256 hash of verification token \\
\texttt{expires\_at} & INTEGER & NOT NULL & Unix timestamp for expiry \\
\texttt{created\_at} & INTEGER & NOT NULL & Unix timestamp when created \\
\bottomrule
\end{tabularx}

\vspace{0.5em}
Constraints and guarantees:
\begin{itemize}[leftmargin=*,nosep]
  \item Primary key on \texttt{user\_id} ensures one token per user.
  \item Only SHA256 hash is stored; plaintext token is sent via email only.
  \item Token is valid only if \texttt{expires\_at > current\_time}.
  \item Token is deleted after successful verification.
  \item Cascading delete removes token when user is deleted.
  % Implementation detail - commented out for Registration_3 draft
  % \item Hourly cleanup task deletes expired tokens and associated unverified users.
\end{itemize}

% Implementation details - commented out for Registration_3 draft
% \subsection{Database Configuration}
%
% \begin{itemize}[leftmargin=*,nosep]
%   \item \textbf{Engine}: SQLite with SQLCipher encryption
%   \item \textbf{Development}: Unencrypted \texttt{data\_dev.db}
%   \item \textbf{Production}: Encrypted \texttt{data.db}
%   \item \textbf{Encryption Key}: 32-byte key (64 hex characters)
%   \begin{itemize}[leftmargin=1.5em,nosep]
%     \item Priority 1: \texttt{APPSEC\_DB\_KEY} environment variable
%     \item Priority 2: System keyring (service: \texttt{APPSEC\_DB\_KEY}, user: \texttt{APPSEC})
%     \item Auto-generated if not found
%   \end{itemize}
%   \item \textbf{Foreign Keys}: Enabled with \texttt{PRAGMA foreign\_keys = ON}
%   \item \textbf{Cleanup}: Hourly task removes expired tokens and unverified users
% \end{itemize}
