% Component Requirements - Functional and non-functional requirements
\section{Component Requirements}

\subsection{Functional Requirements}

\subsubsection{FR-1: User Registration}

\textbf{ID}: FR-1

\textbf{Description}: The system must allow a new user to register with username, email, and password.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Registration is exposed via \texttt{POST /api/register} endpoint.
  \item The frontend provides a registration form with real-time validation.
  \item Username must be 3--20 characters, printable UTF-8 only.
  \item Email must be valid format (validated using \texttt{lettre::Address}).
  \item Password must be 8--64 characters with uppercase, lowercase, digit, and special character.
  \item Returns specific error codes: \texttt{USERNAME\_TAKEN}, \texttt{EMAIL\_TAKEN}, \texttt{VALIDATION}.
\end{itemize}

\subsubsection{FR-2: Email Verification}

\textbf{ID}: FR-2

\textbf{Description}: The system must require email verification before the account is active.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item On successful registration, a 32-byte verification token is generated.
  \item Token is stored as SHA256 hash in \texttt{email\_verification\_tokens} table.
  \item Verification email is sent via SMTP with link: \texttt{\{BASE\_URL\}/verify-email?token=\{token\}}.
  \item Token has configurable expiry (\texttt{EMAIL\_VERIFICATION\_TOKEN\_DURATION\_HOURS}).
  \item \texttt{POST /api/verify-email} endpoint verifies token and marks user as verified.
  \item Token is deleted after successful verification.
\end{itemize}

\subsubsection{FR-3: Input Validation}

\textbf{ID}: FR-3

\textbf{Description}: All registration inputs must be validated on both frontend and backend.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Shared validation logic in \texttt{field-validator} crate (compiled to WASM for frontend).
  \item Username: 3--20 characters, printable UTF-8.
  \item Email: Valid format checked via \texttt{lettre::Address}.
  \item Password: 8--64 characters, at least 1 uppercase, 1 lowercase, 1 digit, 1 special character.
  \item Password strength score calculated (0--7 scale): Weak ($\leq$3), Medium (4--5), Strong (6), CIA (7).
  \item Validation errors are translated to user-friendly messages via \texttt{translator} crate.
\end{itemize}

\subsubsection{FR-4: Error Handling}

\textbf{ID}: FR-4

\textbf{Description}: The system must provide clear, translated error messages.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Error responses use typed error codes from \texttt{api-types} crate.
  \item Validation errors include field-specific error codes (e.g., \texttt{TOO\_SHORT}, \texttt{INVALID\_FORMAT}).
  \item All error codes are translated via \texttt{translator} crate with localization support.
  \item Server errors logged internally; clients receive generic \texttt{INTERNAL} error.
\end{itemize}

\subsubsection{FR-5: Token Lifecycle}

\textbf{ID}: FR-5

\textbf{Description}: Verification tokens must be single-use and time-limited.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Each user can have only one active verification token (primary key on \texttt{user\_id}).
  \item Tokens expire after configurable duration.
  \item Expired tokens are automatically cleaned up by hourly cleanup task.
  \item Unverified users with expired tokens are deleted (cascading foreign key).
  \item Token is deleted upon successful verification.
\end{itemize}

\subsubsection{FR-6: Health Check}

\textbf{ID}: FR-6

\textbf{Description}: The module must expose a health endpoint for monitoring.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item \texttt{GET /api/health} returns basic health status.
  \item Does not expose internal state or user data.
\end{itemize}

\subsection{Non-Functional Requirements}

\subsubsection{NFR-1: Security}

\textbf{ID}: NFR-1

\textbf{Description}: All security best practices must be followed.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Passwords hashed with Argon2 (memory-hard, GPU-resistant).
  \item Verification tokens stored as SHA256 hashes only.
  \item Database encrypted with SQLCipher (32-byte key from keyring or environment).
  \item HTTPS enforced in production.
  \item Input validated on both frontend and backend.
  \item Generic error responses prevent user enumeration.
  \item Foreign keys with CASCADE ensure data cleanup on user deletion.
\end{itemize}

\subsubsection{NFR-2: Performance}

\textbf{ID}: NFR-2

\textbf{Description}: The module must handle typical registration loads efficiently.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item SQLite database with connection pooling.
  \item Argon2 parameters balanced for security and latency.
  \item Hourly cleanup task runs asynchronously.
  \item Frontend validation reduces unnecessary backend requests.
\end{itemize}

\subsubsection{NFR-3: Reliability}

\textbf{ID}: NFR-3

\textbf{Description}: The module must be fault-tolerant.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item On email sending failure, user record is deleted (rollback).
  \item Foreign keys with CASCADE ensure referential integrity.
  \item Automatic cleanup of expired tokens prevents database bloat.
  \item Verification endpoint is idempotent (returns success if already verified).
\end{itemize}

\subsubsection{NFR-4: Usability}

\textbf{ID}: NFR-4

\textbf{Description}: The module must provide clear user feedback.

\textbf{Details}:
\begin{itemize}[leftmargin=*,nosep]
  \item Real-time field validation on frontend.
  \item Password strength indicator with 0--7 score and strength labels.
  \item Translated error messages for all validation failures.
  \item Success message instructs user to check email for verification.
  \item Verification page shows clear success/error states with navigation.
\end{itemize}
