% Security Features Summary - Implemented measures and future enhancements
\section{Security Features Summary}

\subsection{Implemented Security Measures}

\subsubsection*{Password Hashing}

\begin{itemize}[leftmargin=*,nosep]
  \item Argon2 algorithm (memory-hard, GPU-resistant).
  \item Random salt generated using \texttt{SaltString::generate(\&mut OsRng)}.
  \item Password hash stored in PHC string format (includes algorithm, parameters, salt).
  \item Passwords are never logged, stored in plaintext, or returned to clients.
\end{itemize}

\subsubsection*{Token Security}

\begin{itemize}[leftmargin=*,nosep]
  \item Verification tokens are 32 bytes of cryptographically secure random data.
  \item Tokens are hex-encoded for transmission (64 characters).
  \item Only SHA256 hashes of tokens are stored in the database.
  \item Plaintext tokens are sent only via email and never logged.
  \item Tokens have configurable expiry (\texttt{EMAIL\_VERIFICATION\_TOKEN\_DURATION\_HOURS}).
  \item One token per user (primary key constraint).
\end{itemize}

\subsubsection*{Database Encryption}

\begin{itemize}[leftmargin=*,nosep]
  \item SQLite database encrypted with SQLCipher.
  \item 32-byte encryption key (256-bit AES).
  \item Key stored in system keyring or environment variable (\texttt{APPSEC\_DB\_KEY}).
  \item Development mode uses unencrypted database for convenience.
\end{itemize}

\subsubsection*{Input Validation}

\begin{itemize}[leftmargin=*,nosep]
  \item Dual validation: frontend (WASM) and backend (native Rust).
  \item Shared validation logic ensures consistency via \texttt{field-validator} crate.
  \item Username: 3--20 printable UTF-8 characters.
  \item Email: Valid format validated with \texttt{lettre::Address}.
  \item Password: 8--64 characters, requires uppercase, lowercase, digit, special character.
  \item Backend validation is authoritative; frontend validation improves UX.
\end{itemize}

\subsubsection*{Error Handling and Enumeration Resistance}

\begin{itemize}[leftmargin=*,nosep]
  \item Typed error codes from \texttt{api-types} crate.
  \item Specific codes for \texttt{USERNAME\_TAKEN} and \texttt{EMAIL\_TAKEN} (HTTP 409).
  \item Validation errors include field-specific details for user feedback.
  \item Internal errors return generic \texttt{INTERNAL} code (HTTP 500).
  \item All error codes translated to user-friendly messages via \texttt{translator} crate.
\end{itemize}

\subsubsection*{Automatic Cleanup}

\begin{itemize}[leftmargin=*,nosep]
  \item Hourly cleanup task runs asynchronously.
  \item Deletes expired verification tokens.
  \item Deletes unverified users whose tokens have expired (via CASCADE).
  \item Prevents database bloat from abandoned registrations.
\end{itemize}

\subsubsection*{Email Delivery}

\begin{itemize}[leftmargin=*,nosep]
  \item SMTP-based email sending via \texttt{lettre} crate.
  \item Development: MailHog at \texttt{127.0.0.1:1025}.
  \item Production: Configurable SMTP server.
  \item On email failure, user record is deleted (transactional rollback).
\end{itemize}

\subsubsection*{Foreign Key Integrity}

\begin{itemize}[leftmargin=*,nosep]
  \item Foreign keys enabled with \texttt{PRAGMA foreign\_keys = ON}.
  \item CASCADE delete on all child tables.
  \item Ensures data consistency when users are deleted.
\end{itemize}

\subsection{Future Security Enhancements}

Planned or potential enhancements (Registration\_3 scope):

\begin{itemize}[leftmargin=*,nosep]
  \item Rate limiting on registration endpoint.
  \item CAPTCHA integration for bot mitigation.
  \item Email domain whitelist/blacklist.
  \item Invite token system for controlled registration.
  % Login_4.md related - commented out for Registration_3.md draft
  % \item Session management using secure cookies.
  % \item Password reset flow with secure tokens.
  % \item Account lockout after repeated failures.
  % \item Multi-factor authentication (MFA).
\end{itemize}
